\mythesistitle

### 2011-12-01 src/synopsis.tex
final version before spell check

% \begin{figure}[t]%[htbp]
%   \centering
%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
%   \caption{{\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%   \label{fig:streetmappath}
% \end{figure}
% \begin{figure}[t] % [h]
%   \centering
%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
%   \caption{Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}
%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
% \begin{figure}[htbp] %[t]%
%  \centering
%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
%  \label{fig:streetmap}
%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
% \end{figure}
\begin{figure}[htbp] %[t]%
  \centering
  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
  \caption{\figtabsize Individual allocation of apartments to students
    in {\residenceblock} that meets the requirements stated before.
    The routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green
    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images {\copyright} Charles
    Schulz}}}
  \label{fig:streetmappathpeople}
\end{figure}

@@

Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it.  Section~\ref{sec:results}
presents a summary of our results on the extention of COP namely, the
tree path labeling problem. \tnote[TH]{update for thesis}

@@

@@

The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  A landmark result came a few years later
when \cite{at72} discovered the families of forbidden submatrices that
prevent a matrix from having COP and most, if not all, results that
came later were based on this discovery which connected COP in binary
matrices to convex bipartite graphs. In fact, the forbidden
submatrices came as a corollary to the discovery that convex bipartite
graphs are AT-free in \cite{at72}\tnote{check}. The first linear time
algorithm for COP testing (COT) was invented by \cite{bl76} using a
data structure called PQ trees.  Since then several COT algorithms
have been invented -- some of which involved variations of PQ trees
\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
certifying algorithms\cite{mcc04}. % These algorithms
% improve over PQ trees on the time complexity and/or the complexity
% of implementation.

The construction of PQ trees in \cite{bl76} draws on the
close relationship of matrices with COP to interval graphs. A PQ tree of a
matrix is one that stores all row (column) permutations of the matrix
that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
is constructed using an elaborate linear time procedure and is also
a test for planarity\tnote{check check check. both interval
  graph and planarity in this paper?}. % Since this
% result, several improvements have been made to the procedural
% complexity of the algorithm itself and also derivative data structures
% that are easier to construct.
PQR trees is a generalized data structure based on PQ trees
\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
describes the simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}

@@

\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
  surveyed. hardness results, approx. results. results are usually for
  a class of matrices $(a,b)$ where number {\un}s in columns and rows
  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
  entries of $M$ to make it attain COP. this is NP complete
  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
  are AT-free.  describe AT (2) show the close relationship b/w COP
  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
does not have COP have certain substructures that prevent it from
having COP. Tucker classified these forbidden substructures into five
classes of submatrices. This result is presented in the context of
convex bipartite graphs which \cite{at72} proved to be
AT-free\tnote[TH]{ check this up. give details. - doms'}. By
definition, convex bipartite graph have half adjacency matrices that
have COP on either rows or columns (graph is biconvex if it has COP on
both)\cite{d08phd}. A half adjacency matrix is a binary matrix
representing a bipartite graph as follows. The set of rows and the set
of columns form the two partitions of the graph. Each row node is
adjacent to those nodes that represent the columns that have ones in
the corresponding row. \cite{at72} proves that this bipartite graph
has no asteroidal triple if and only if the matrix has COP and goes on
to identify the forbidden substructures for these bipartite
graphs. The matrices corresponding to these substructures are the
forbidden submatrices.

Once a matrix has been detected to not have COP (using any of the COT
algorithms mentioned earlier), it is naturally of interest to find out
the smallest (in terms of number of rows and/or columns and/or number
of entries that are {\un}s) forbidden substructure. \cite{d08phd}
discusses a couple of algorithms which are efficient if the number of
ones in a row is small. This is of significance in the case of sparce
matrices where this number is much lesser than the number of
columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
on number of {\un}s in any column but has at most $\Delta$ {\un}s in
any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
similar problems which deals with inducing COP on a matrix. In {\sc
  Min COS-R (Min COS-C)} the question is to find the minimum number of
rows (columns) that must be deleted to result in a matrix with COP.
In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
maximum number of rows (columns) that induces a submatrix with
COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
a submatrix $M'$ with all columns\tnote{check if b75 deals with COP
  col or COP row. also is it any submatrix with k less than r rows or
  submatrix must have all columns?} but a maximum cardinality subset
of rows such that $M'$ has COP is NP complete. \cite{hg02} corrects an
error of the abridged proof of this reduction as given in \cite{gj79}.
\cite{d08phd} discusses all these problems in detail giving an
extensive survey of the previously existing results which are almost
exhaustively all approximation results and hardness results. Taking
this further, \cite{d08phd} presents new results in the area of
parameterized algorithms for this problem\tnote[TH]{elaborate - what
  are the results?}.

Another problem is to find the minimum number of entries in the matrix
that can be toggled to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to {\un}s so that the
resulting matrix has COP. As mentioned earlier, this problem is
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP (instead
of one block of consecutive {\un}s, they have $x$ blocks, or {\em
  runs}, of consecutive {\un}s and $x$ is not too large) in the
storage of very large databases.  The problem is that of reordering of
a binary matrix such that the resulting matrix has at most $k$ runs of
{\un}s. This is proven to be NP hard using a reduction from the
hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
\tnote[TH]{(1) A connection of COP problem to the travelling salesman
  problem is also introduced. what does this mean? -- COP can be used
  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
version of the $k$-run problem, i.e. minimization of number of blocks
of ones is proven to be NP complete by cite:k77}\tnote{are these two
  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

@@

We give a necessary and sufficient condition (called ICPPL)\tnote{say a few
  words about icppl} for {\FTPL} to output in
affirmative. This characterization can be checked in polynomial time.
The most interesting consequence is that in our constructive
procedure, it is sufficient to iteratively check if three-way
intersection cardinalities are preserved.  In other words, in each
iteration, it is sufficient to check if the intersection of any three
hyperedges is of the same cardinality as the intersection of the
corresponding paths.  Thus this generalizes the well studied question
of the feasible interval assignment problem which is the special case
when the target tree $T$ is simply a path \cite{wlh02,nsnrs09}.

@@

We also have an algorithm for solving {\CFTPL} with no restrictions
which runs in exponential time and explores a partial order theory of
matrices related to {FTPL}. This po theory gives interesting insight
into the problem of finding path labeling from $T$ by decomposing it
into smaller subproblems of finding path labeling of subsets
of $\cF$ from subtrees of $T$.\tnote[TH]{The connection
of TPL to graph isomorphism will be made later in the document}
%\end{enumerate}

@@

\section{Conclusion}

Our results also have close connection to recognition of path graphs
and connection to path graph isomorphism.  Graphs which can be
represented as the intersection graph of paths in a tree are called
{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
be viewed as paths in a tree, if and only if the intersection graph of
$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
graphs since chordal graphs are combinatorially characterized as the
intersection graphs of subtrees of a tree.  Path graphs are well
studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
a graph is a path graph can be done in polynomial time by the results
of \cite{gav78,aas93}.  However, this is only a necessary condition
for our question.  \remove{Indeed, our work provides a sufficient
  condition for a related problem: that is whether a path assignment
  to be feasible?}  On the other hand, path graph isomorphism is known
be isomorphism-complete; see for example \cite{kklv10}. Therefore,
\annote{it is unlikely that we can solve the problem of finding
  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
\annote{It is definitely interesting to classify the kinds of trees
  and hypergraphs for which feasible path labelings can be
}{ELABORATE.} found efficiently.  \annote{These results would form
  a natural generalization of COP testing and interval graph
  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.

\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[TH]{peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the

ordering\cite{gh64}. This clearly means that a graph $G$ is an
interval graph if and only if \tnote{verify from paper the
  statement of claim.} maximal clique vertex incidence matrix of $G$
has COP.  Also, maximal cliques of any chordal graph can be enumerated
in polytime $O(m+n)$ \tnote{citation?!!}.  \cite{fg65} uses these
results to give the first polynomial time algorithm for
COT. \tnote[TH]{check. how do they use it?}  A bipartite graph
is convex (on $R$) if and only if its half adjacency matrix has COP on
rows.  The results in \cite{bl76}
%\cite{bl62} 
on COT are based on the result that
interval graphs are AT-free chordal graphs. \tnote{the latter being
  Tucker's?}

\tnote[TH]{(1) How COP (and TPL) relates to interval graph isomorphism
  (and path graph isomorphism) move/summarize what is already in
  Motivation section.(2) TBD survey -- see blue notes under Graph
  Isomorphism. namely citations in cite:aas93 (3) a brief on heirarchy
  of chordal graphs, path graphs, interval graphs, peo, clique tree
  etc. and results. (4) interval graphs are incomparability
  graphs. see Golumbic. [tag:classification] (5) finding min length
  hole in bipartite graph is polynomial time [Sec 3.3.2 in
  cite:d08phd] (6) theorem 2.2 in cite:d08phd - $G$ is union of
  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
  incidence matrix $(*,2)$ of $G$}

{\bf Complexity challenges:} Canonization is an important tool in graph isomorphism. Invention of a
deterministic method of canonization for any class of graphs naturally
results in an algorithm for isomorphism. All that is required is to
check if the canons of two graphs are the same. Thus complexity of
graph isomorphism can be studied by studying canonization
methods. While general graph ismorphism remains elusive in terms of
hardness, canonization has been studied for smaller classes of graphs
thus giving complexity/hardness results for them.

In 1992, \cite{sl92} made an important discovery that tree isomorphism
is in logspace. It was done by inventing a method of canonization of
trees using a logspace depth first traversal
algorithm. \tnote{confirm the statement of claim}.

\cite{kklv10} proved that interval graph canonization is also in
logspace thus drawing logspace conslusions about COT. 
% This result is
% built on the fact that 
Interval graphs are FP+C (fixed point with
counting) definable \tnote{elucidate} and \cite{l10} showed that
this implies that it captures PTIME. \tnote{elucidate} This result
along with that of undirected graph reachability being in logspace
\cite{rei08}, \cite{kklv10} proved their logspace result.\\

\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
  (2) [[[COP helping some problems with hardness]]] having COP in the
  input structure makes some problems less hard than in a general
  input. -- min set cover problem is known to be NP hard. so is min
  hitting
  set. these two problems are equivalent cite-adp80\\
  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
  -- a couple of generalizations:\\
  -- -- min degree hypergraph (MDH) problem\\
  -- -- red blue covering (RBSC) problem\\
  ----- cite-d08phd Chapter 5.\\
  ----- what are cite-d08phd results\\
  ----- Rectangle stabbing \\
  ----- see cite-d08phd Chapter 6?  }

@@

\item Research %- Extension to trees
  \begin{enumerate}[{Section 4.}1]
  \item Tree path labeling of path
    hypergraphs%: ( stuff in latin paper,)
  \item Extension of po theory from \cite{nsnrs09}
    % (the matrix theory submitted in walcom, any other work - need to
    % see notes.)
  \end{enumerate}

>>>>

% \begin{figure}[t]%[htbp]
%   \centering
%   \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
%   \caption{{\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
%   \label{fig:streetmappath}
% \end{figure}
% \begin{figure}[t] % [h]
%   \centering
%   \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
%   \caption{Individual allocation of apartments to students in
%     {\residenceblock} that meets the requirements stated before.}
%   \label{fig:streetmappathpeople}
% \end{figure}
%%%%%%%%%%%% END THESIS UNCOMMENT %%%%%%%%%%%%
% \begin{figure}[htbp] %[t]%
%  \centering
%   (a) \includegraphics[scale=0.2]{../img/1_infinite_loop.png}
%   (b) \includegraphics[scale=0.2]{../img/2_infinite_loop_BTWF.png}
%  \label{fig:streetmap}
%   \caption{\figtabsize (a) {\residenceblock} street map (b) {\residenceblock} street map with study group routes
%     allocated. Routes are color coded as follows: red for
%     \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
%     group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
%     \textcolor{green}{$\xTTT$} group}
% \end{figure}
\begin{figure}[htbp] %[t]%
  \centering
  \includegraphics[scale=0.3]{../img/3_infinite_loop.png}
  \caption{\figtabsize Individual allocation of apartments to students
    in {\residenceblock} that meets the requirements stated before.
    The routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green
    for \textcolor{green}{$\xTTT$} group. {\tiny {\em Peanuts images
        {\copyright} Charles Schulz}}}
  \label{fig:streetmappathpeople}
\end{figure}
\tnote[TH]{make dashed/textured lines for routes. make it color agnostic.}

@@

Section~\ref{sec:background} gives a brief survey of COP and
optimization problems related to it followed by motivation for the
thesis in Section~\ref{sec:motive}.  Section~\ref{sec:results}
presents a summary of our results on the extention of COP namely, the
tree path labeling problem. \tnote[TH]{update for thesis}

@@

%%%%%%                              %%%%%%
%THX is the second level of ``tnote''-ing%
%%%%%%                              %%%%%%
\tnote[TH]{ADD: As it will be described in detail later in this
  document, isomorphism of certain % classes of graphs, namely chordal
%   graphs, have a close relationship with consecutive ones property and
%   generalizations of it.  This is perhaps because of how closely COP
%   of a matrix relates to properties of graphs derived from matrices as
%   seen in the following results.
}
\tnote[THX]{ADD: peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} %
\tnote[THX]{ADD: A well known result in %perfect graph theory is that
%  the maximal cliques of  an interval graph $G$ can be linearly ordered
%   such that for all $v \in V(G)$, cliques containing $v$ are
%   consecutive in the ordering cite~{gh64}. This clearly means that a
%   graph $G$ is an interval graph if and only if
}% 
\tnote[THX]{verify
  from paper the statement of claim.}% 
\tnote[TH]{maximal clique vertex
  incidence matrix of%  $G$ has COP.  Also, maximal cliques of any
%   chordal graph can be enumerated in polytime $O(m+n)$
}
\tnote[THX]{citation?!!}%  
\tnote[TH]{cite~{fg65} uses these results to
  give the first polynomial time algorithm for COT.}%
\tnote[THX]{check. how do they use it?}%  
\tnote[TH]{ A bipartite graph
  is convex%  (on $R$) if and only if its half adjacency matrix has COP
%   on rows.  The results in cite~{bl76} on COT are based on the result
%   that interval graphs are AT-free chordal graphs.
}%
\tnote[THX]{the latter being Tucker's?}%  
\tnote[THX]{(2) TBD survey
  -- % see blue notes (in notebook) under Graph Isomorphism. namely
%   citations in cite:aas93 (3) a brief on heirarchy of chordal graphs,
%   path graphs, interval graphs, peo, clique tree etc. and results. (4)
%   interval graphs are incomparability graphs. see
%   Golumbic. [tag:classification] (5) finding min length hole in
%   bipartite graph is polynomial time [Sec 3.3.2 in cite:d08phd] (6)
%   theorem 2.2 in cite:d08phd - $G$ is union of v.d. caterpillars iff
%   $M$ has COP in rows. $M$ is edge vertex incidence matrix $(*,2)$ of
%   $G$
}%

@@

The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  A landmark result came a few years later
when \cite{at72} discovered the families of forbidden submatrices that
prevent a matrix from having COP and most, if not all, results that
came later were based on this discovery which connected COP in binary
matrices to convex bipartite graphs. In fact, the forbidden
submatrices came as a corollary to the discovery that convex bipartite
graphs are AT-free in \cite{at72}\tnote[GTC]{check}. The first linear time
algorithm for COP testing (COT) was invented by \cite{bl76} using a
data structure called PQ trees.  Since then several COT algorithms
have been invented -- some of which involved variations of PQ trees
\cite{mm96,wlh01,mcc04}, some involved set theory and ICPIA
\cite{wlh02,nsnrs09}, parallel COT algorithms\cite{as95,bs03,ly91} and
certifying algorithms\cite{mcc04}. % These algorithms
% improve over PQ trees on the time complexity and/or the complexity
% of implementation.

The construction of PQ trees in \cite{bl76} draws on the
close relationship of matrices with COP to interval graphs. A PQ tree of a
matrix is one that stores all row (column) permutations of the matrix
that give the COP orders (there could be multiple orders of rows or columns) of the matrix. This
is constructed using an elaborate linear time procedure and is also
a test for planarity\tnote[GTC]{check check check. both interval
  graph and planarity in this paper?}. % Since this
% result, several improvements have been made to the procedural
% complexity of the algorithm itself and also derivative data structures
% that are easier to construct.
PQR trees is a generalized data structure based on PQ trees
\cite{mm96,mpt98}.  \cite{tm05} describes an improved algorithm to
build PQR trees. \tnote[GTC]{improv in terms of what?}\cite{wlh02}
describes the simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote[GTC]{it reduces the solution search space. fill in the blanks.}

@@

\tnote[TH]{ -- sect 4.1 in cite:d08phd has many results
  surveyed. hardness results, approx. results. results are usually for
  a class of matrices $(a,b)$ where number {\un}s in columns and rows
  are restriced to $a$ and $b$ . -- problem of flipping at most $k$
  entries of $M$ to make it attain COP. this is NP complete
  cite:b75-phd}\tnote[TH]{(1) scite:lb62 showed that interval graphs
  are AT-free.  describe AT (2) show the close relationship b/w COP
  and graphs sec 2.2, pg 31} \cite{at72} showed that a matrix that
does not have COP have certain substructures that prevent it from
having COP. Tucker classified these forbidden substructures into five
classes of submatrices. This result is presented in the context of
convex bipartite graphs which \cite{at72} proved to be
AT-free\tnote[TH]{ check this up. give details. - doms'}. By
definition, convex bipartite graph have half adjacency matrices that
have COP on either rows or columns (graph is biconvex if it has COP on
both)\cite{d08phd}. A half adjacency matrix is a binary matrix
representing a bipartite graph as follows. The set of rows and the set
of columns form the two partitions of the graph. Each row node is
adjacent to those nodes that represent the columns that have {\un}s in
the corresponding row. \cite{at72} proves that this bipartite graph
has no asteroidal triple if and only if the matrix has COP and goes on
to identify the forbidden substructures for these bipartite
graphs. The matrices corresponding to these substructures are the
forbidden submatrices.

Once a matrix has been detected to not have COP (using any of the COT
algorithms mentioned earlier), it is naturally of interest to find out
the smallest forbidden substructure (in terms of number of rows and/or
columns and/or number of entries that are {\un}s). \cite{d08phd}
discusses a couple of algorithms which are efficient if the number of
{\un}s in a row is small. This is of significance in the case of sparce
matrices where this number is much lesser than the number of
columns. $(*,\Delta)${\em -matrices} are matrices with no restriction
on number of {\un}s in any column but has at most $\Delta$ {\un}s in
any row. {\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are
similar problems which deals with inducing COP on a matrix. In {\sc
  Min COS-R (Min COS-C)} the question is to find the minimum number of
rows (columns) that must be deleted to result in a matrix with COP.
In the dual problem {\sc Max COS-R (Max COS-C)} the search is for the
maximum number of rows (columns) that induces a submatrix with
COP. Given a matrix $M$ with no COP, \cite{b75-phd} shows that finding
a submatrix $M'$ with all columns\tnote[GTC]{check if b75 deals with
  COP col or COP row. also is it any submatrix with k less than r rows
  or submatrix must have all columns?} but a maximum cardinality
subset of rows such that $M'$ has COP is NP complete. \cite{hg02}
corrects an error of the abridged proof of this reduction as given in
\cite{gj79}.  \cite{d08phd} discusses all these problems in detail
giving an extensive survey of the previously existing results which
are almost exhaustively all approximation results and hardness
results. Taking this further, \cite{d08phd} presents new results in
the area of parameterized algorithms for this
problem\tnote[TH]{elaborate - what are the results?}.

Another problem is to find the minimum number of entries in the matrix
that can be toggled to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to {\un}s so that the
resulting matrix has COP. As mentioned earlier, this problem is
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote[GTC]{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote[GTC]{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP (instead
of one block of consecutive {\un}s, they have $x$ blocks, or {\em
  runs}, of consecutive {\un}s and $x$ is not too large) in the
storage of very large databases.  The problem is that of reordering of
a binary matrix such that the resulting matrix has at most $k$ runs of
{\un}s. This is proven to be NP hard using a reduction from the
hamiltonian path problem.\tnote[TH]{Theorem 2.1 in jkckv}
\tnote[TH]{(1) A connection of COP problem to the travelling salesman
  problem is also introduced. what does this mean? -- COP can be used
  as a tool to reorder $0.5T \le runs(M) \le T$. (2) The optimization
version of the $k$-run problem, i.e. minimization of number of blocks
of ones is proven to be NP complete by cite:k77}\tnote[GTC]{are these two
  the same?} \tnote[TH]{what is the reduction?} \tnote[TH]{other
  problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) --
  cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

@@

We give a necessary and sufficient condition (called ICPPL)\tnote[GTC]{say
  a few words about icppl} for {\FTPL} to output in affirmative. This
characterization can be checked in polynomial time.  The most
interesting consequence is that in our constructive procedure, it is
sufficient to iteratively check if three-way intersection
cardinalities are preserved.  In other words, in each iteration, it is
sufficient to check if the intersection of any three hyperedges is of
the same cardinality as the intersection of the corresponding paths.
Thus this generalizes the well studied question of the feasible
interval assignment problem which is the special case when the target
tree $T$ is simply a path \cite{wlh02,nsnrs09}.

@@

We also have an algorithm for solving {\CFTPL} with no restrictions
which runs in exponential time. 
% and explores a partial order theory of
% matrices related to {FTPL}.
% gives interesting insight
% into the problem of 
This algorithm finds a path labeling from $T$ by decomposing the
problem into smaller subproblems of finding path labeling of subsets
of $\cF$ from subtrees of $T$. Given the fact that binary matrices
naturally represent a set
system % , we look at the matrix representation of subsets of $\cF$.
and that the overlap relation between the sets involved is an obvious
equivalence relation, $\cF$ naturally partitions into equivalence
classes known as overlap components. In the context of COP, {\em overlap
components} were used in \cite{wlh02} and \cite{kklv10}. Moreoever,
\cite{nsnrs09} discovered that these equivalence classes form a total
order. We extend this to TPL and find that when $\cF$ is a path
hypergraph\footnote{If there exists an FTPL for a hypergraph $\cF$, it
  is called a path hypergraph.}, the classes can be partially ordered
as an in-tree in polynomial time. Once $\cF$ is
``broken'' into overlap components, one must identify the subtree of
$T$ that it needs to map to and this is the hard part which is
currently open to be solved in polynomial time.

\tnote[TH]{The connection of TPL to graph
  isomorphism will be made later in the document}

@@

\section{Conclusion}
\label{sec:conclusion}
\vspace{\secfirstparatrim}

We give a characterization for feasible tree path labeling of path
hypergraphs. The proof for this is constructive and computes a
feasiblility bijection mapping vertices of the hypergraph to vertices
of the given target tree. This thesis also discovered an exponential algorithm
that computes a feasible TPL to a given hypergraph if it is a path hypergraph.

Our results have close connections to recognition of path graphs and
to path graph isomorphism.  Graphs which can be represented as the
intersection graph of paths in a tree are called {\em path
  graphs}\cite{mcg04}. Thus, a hypergraph $\cF$ can be interpreted as
paths in a tree, if and only if the intersection graph of $\cF$ is a
{\em path graph}. Path graphs are a subclass of chordal graphs since
chordal graphs are characterized as the intersection graphs of
subtrees of a tree\cite{mcg04}.  Path graphs are well studied in the
literature \cite{plr70}--\cite{mcg04}\tnote[TH]{Check if cite:bp93 is
  indeed for path graphs}.  Path graph recognition can be done in
polynomial time\cite{gav78,aas93}.  Clearly, this is a necessary
condition in terms of the intersection graph of the input hypergraph
$\cF$ in \FTPL. However, one can easily obtain a counterexample to
show the insufficiency of this condition\tnote[GTC]{Give a figure
  showing insuff.}.  Path graph isomorphism is known be
isomorphism-complete\cite{kklv10}. Therefore, \annote[GTC]{it is unlikely
  that we can solve the problem of finding feasible path labeling
  $\cl$ for a given $\cF$ and }{WHY?} tree $T$.  It is
  definitely interesting to classify the kinds of trees and
  hypergraphs for which feasible path labelings can be\tnote[TH]{ELABORATE.}
found efficiently.  \annote[TH]{These results would form a natural
  generalization of COP testing and interval graph isomorphism,
  culminating in Graph Isomorphism }{REWORD} itself.
 To this effect we consider TPL on {\kstar} and give a polynomial time
 solution for the same.

Whether TPL on general trees is solvable in $P$ remains open. So do
optimization opportunities in TPL (possible extentions to optimization for COP in the
Section~\ref{sec:optcop}.\tnote[TH]{link to section}).

\tnote[TH]{ ||Complexity challenges:|| Canonization is an important tool in graph isomorphism. Invention of a
% deterministic method of canonization for any class of graphs naturally
% results in an algorithm for isomorphism. All that is required is to
% check if the canons of two graphs are the same. Thus complexity of
% graph isomorphism can be studied by studying canonization
% methods. While general graph ismorphism remains elusive in terms of
% hardness, canonization has been studied for smaller classes of graphs
% thus giving complexity/hardness results for them.

% In 1992, \cite{sl92} made an important discovery that tree isomorphism
% is in logspace. It was done by inventing a method of canonization of
% trees using a logspace depth first traversal
% algorithm. \tnote{confirm the statement of claim}.

% \cite{kklv10} proved that interval graph canonization is also in
% logspace thus drawing logspace conslusions about COT. 
% % This result is
% % built on the fact that 
% Interval graphs are FP+C (fixed point with
% counting) definable \tnote{elucidate} and \cite{l10} showed that
% this implies that it captures PTIME. \tnote{elucidate} This result
% along with that of undirected graph reachability being in logspace
% \cite{rei08}, \cite{kklv10} proved their logspace result.
}%
\tnote[TH]{ (1) the interesting survey in conclusion section of kklv10
  (2) [[[COP helping some problems with hardness]]] having COP in the
  input structure makes some problems less hard than in a general
  input. -- min set cover problem is known to be NP hard. so is min
  hitting
  set. these two problems are equivalent cite-adp80\\
  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
  -- a couple of generalizations:\\
  -- -- min degree hypergraph (MDH) problem\\
  -- -- red blue covering (RBSC) problem\\
  ----- cite-d08phd Chapter 5.\\
  ----- what are cite-d08phd results\\
  ----- Rectangle stabbing \\
  ----- see cite-d08phd Chapter 6?  }

@@

\item Research %- Extension to trees
  \begin{enumerate}[{Section 4.}1]
  \item Tree path labeling of path
    hypergraphs %: ( stuff in latin paper,)
  \item Extension of the structural characterization of matrices in \cite{nsnrs09}
    % (the matrix theory submitted in walcom, any other work - need to
    % see notes.)
  \end{enumerate}
